---
title: "MT Cars Investigation"
author: "Andrew Fuller | AROC Data Scientist"
format: html
editor: source
---

```{r}
#| label: setup
#| echo: false
library(datasets)
```

## Overview

The `mtcars` dataset comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

```{r}
#| label: glimpse
str(mtcars)
```

## Summary Statistics

```{r}
#| label: summary-stats
summary(mtcars)
```

## MPG by Cylinder Count

```{r}
#| label: mpg-by-cyl
boxplot(mpg ~ cyl, data = mtcars,
        xlab = "Cylinders", ylab = "Miles per Gallon",
        main = "MPG by Number of Cylinders",
        col = c("steelblue", "orange", "tomato"))
```

## Horsepower vs MPG

```{r}
#| label: hp-vs-mpg
plot(mtcars$hp, mtcars$mpg,
     xlab = "Horsepower", ylab = "Miles per Gallon",
     main = "Horsepower vs MPG",
     pch = 19, col = "steelblue")
abline(lm(mpg ~ hp, data = mtcars), col = "tomato", lwd = 2)
```

Cars with higher horsepower tend to have lower fuel efficiency (MPG), as shown by the downward trend.

## Weight vs MPG

```{r}
#| label: wt-vs-mpg
plot(mtcars$wt, mtcars$mpg,
     xlab = "Weight (1000 lbs)", ylab = "Miles per Gallon",
     main = "Weight vs MPG",
     pch = 19, col = "forestgreen")
abline(lm(mpg ~ wt, data = mtcars), col = "tomato", lwd = 2)
```

Heavier cars show a strong negative correlation with fuel efficiency.
