---
title: "MT Cars Investigation"
author: "Andrew Fuller | AROC Data Scientist"
format: html
editor: source
params:
  cyl:
    value: "all"
    input: select
    label: "Filter by cylinders"
    choices: ["all", "4", "6", "8"]
  am:
    value: "all"
    input: select
    label: "Filter by transmission"
    choices: ["all", "0", "1"]
  show_plots:
    value: true
    input: checkbox
    label: "Show plots"
  title_suffix:
    value: ""
    input: text
    label: "Title suffix"
---

```{r}
#| label: setup
#| echo: false
library(datasets)

mt <- mtcars

filter_msg <- character()

if (params$cyl != "all") {
  cyl_val <- as.numeric(params$cyl)
  mt <- mt[mt$cyl == cyl_val, ]
  filter_msg <- c(filter_msg, paste0("cyl = ", cyl_val))
}

if (params$am != "all") {
  am_val <- as.numeric(params$am)
  mt <- mt[mt$am == am_val, ]
  filter_msg <- c(filter_msg, paste0("am = ", am_val))
}

filter_label <- if (length(filter_msg) > 0) {
  paste0(" (filtered: ", paste(filter_msg, collapse = ", "), ")")
} else {
  ""
}

data_summary <- if (nrow(mt) < nrow(mtcars)) {
  paste0(" After filtering, ", nrow(mt), " of ", nrow(mtcars), " cars remain.")
} else {
  ""
}
```

## Overview{`r params$title_suffix`}

The `mtcars` dataset comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).`r data_summary`

```{r}
#| label: glimpse
str(mt)
```

## Summary Statistics{`r filter_label`}

```{r}
#| label: summary-stats
summary(mt)
```

## MPG by Cylinder Count{`r filter_label`}

```{r}
#| label: mpg-by-cyl
if (params$show_plots && nrow(mt) > 0) {
  if (length(unique(mt$cyl)) > 1) {
    boxplot(mpg ~ cyl, data = mt,
            xlab = "Cylinders", ylab = "Miles per Gallon",
            main = "MPG by Number of Cylinders",
            col = c("steelblue", "orange", "tomato"))
  } else {
    boxplot(mpg ~ cyl, data = mt,
            xlab = "Cylinders", ylab = "Miles per Gallon",
            main = paste("MPG | Cylinders =", unique(mt$cyl)),
            col = "steelblue")
  }
}
```

## Horsepower vs MPG{`r filter_label`}

```{r}
#| label: hp-vs-mpg
if (params$show_plots && nrow(mt) > 0) {
  plot(mt$hp, mt$mpg,
       xlab = "Horsepower", ylab = "Miles per Gallon",
       main = "Horsepower vs MPG",
       pch = 19, col = "steelblue")
  abline(lm(mpg ~ hp, data = mt), col = "tomato", lwd = 2)
}
```

`r if (nrow(mt) > 0) "Cars with higher horsepower tend to have lower fuel efficiency (MPG), as shown by the downward trend."`

## Weight vs MPG{`r filter_label`}

```{r}
#| label: wt-vs-mpg
if (params$show_plots && nrow(mt) > 0) {
  plot(mt$wt, mt$mpg,
       xlab = "Weight (1000 lbs)", ylab = "Miles per Gallon",
       main = "Weight vs MPG",
       pch = 19, col = "forestgreen")
  abline(lm(mpg ~ wt, data = mt), col = "tomato", lwd = 2)
}
```

`r if (nrow(mt) > 0) "Heavier cars show a strong negative correlation with fuel efficiency."`
